---
title: "p8105_hw3_NL2836"
author: "Ngan Le"
date: "2023-10-12"
output: github_document
---

Load the necessary libraries and datasets.
```{r setup, message = FALSE}
library(tidyverse)
library(p8105.datasets)
library(gt)
library(patchwork)
data("instacart")
data("brfss_smart2010")
```

# Problem 1

##### How many aisles are there, and which aisles are the most items ordered from?

Count the number of distinct aisles.
```{r aisle count}
aisle_count = 
instacart %>% 
  summarize(n_distinct(aisle))
```

Arrange the `aisle` variable according to observation frequencies and output the first (highest frequency) row.
```{r aisle frequency count}
aisle_most_order =
  instacart %>% 
  group_by(aisle) %>% 
  summarize(freq_order = n()) %>% 
  arrange(desc(freq_order)) %>% 
  pull(aisle) %>% 
  first()
```

Answer: There are `r aisle_count` different aisles. The aisle from which most items are ordered is `r aisle_most_order`.

##### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

Filter aisles to those with more than 10000 and arrange.

```{r filter aisle}
aisle_plot =
  instacart %>% 
  group_by(aisle) %>% 
  summarize(freq_aisle= n()) %>%
  filter(freq_aisle > 10000)
```

Make a plot to show the number of items ordered in each aisle.

```{r, dpi = 300}
aisle_plot %>% 
  ggplot(aes(x= reorder(aisle, freq_aisle), y = freq_aisle)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  coord_flip() +
  labs(x = "Aisle name", y = "Number of Items Ordered") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 7)) +
  ggtitle("Number of Items Ordered by Aisle")
```

##### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

Get the 3 most popular items of each aisle.

```{r popular items}
popular_items =
  instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(freq_pop=n(),.groups = 'drop') %>% 
  arrange(aisle, desc(freq_pop)) %>% 
  group_by(aisle)%>%
  top_n(3, wt = freq_pop)
```

Make a table. 

```{r popular items table, warning = FALSE}
popular_items %>%
  gt() %>%
  cols_label(
    aisle = "Aisle",
    product_name = "Product name",
    freq_pop = "Times ordered") %>% 
  tab_header(title = "Top 3 Popular Items by Aisle") %>%
  fmt_number(columns = vars(freq_pop), decimals = 0) 
```

##### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r order time}
time_order =
  instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = round(mean(order_hour_of_day, na.rm = TRUE),2), .groups = 'drop') %>% 
 pivot_wider(
    names_from = order_dow,
    values_from = mean_hour) %>% 
 rename(
   Monday = "0", 
   Tuesday = "1", 
   Wednesday = "2", 
   Thursday = "3", 
   Friday = "4", 
   Saturday = "5", 
   Sunday = "6"
       )
```

```{r order time table, warning = FALSE}
time_order %>%
  gt() %>%
  cols_label(
    product_name = "Product name") %>% 
  tab_header(title = "Mean time order by week date") %>%
  fmt_number(columns = vars(product_name), decimals = 0) 
```

# Problem 2

##### Data cleaning

```{r brfss cleaning}
p2_clean = 
  brfss_smart2010 %>%  
  janitor::clean_names() %>% 
  rename(
    location_abbr = locationabbr,
    location_desc = locationdesc,
    resp_id = respid,
        ) %>% 
  filter(
    topic == "Overall Health",
    response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")
         ) %>% 
    mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), ordered = TRUE))
```

##### In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r 2002}
df2002 =
p2_clean %>% 
  filter(year == "2002") %>% 
  group_by(location_abbr) %>% 
  summarize(dist_2002 = n_distinct(location_desc)) %>% 
  filter(dist_2002 >= 7) %>% 
  pull(location_abbr)
```

In 2002, the states that were observed at 7 or more locations are `r df2002`.

```{r 2010}
df2010 =
p2_clean %>% 
  filter(year == "2010") %>% 
  group_by(location_abbr) %>% 
  summarize(dist_2010 = n_distinct(location_desc)) %>% 
  filter(dist_2010 >= 7) %>% 
  pull(location_abbr)
```

In 2010, the states that were observed at 7 or more locations are `r df2010`.

##### Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

Make the dataframe.

```{r make excellent df}
excellent =
  p2_clean %>% 
  filter(response == "Excellent") %>% 
  select(year, location_abbr, data_value) %>% 
  group_by(year, location_abbr) %>% 
  summarize(mean_value = round(mean(data_value, na.rm = TRUE), 2),.groups = 'drop')
```

Plot the average value.

```{r spaghetti}
ggplot(excellent, aes(x = year, y = mean_value, group = location_abbr, color = location_abbr)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Average Value",
    title = "Spaghetti Plot of Average Value Over Time by State"
  )
```

The plot shows the trend of the average data value of each states from 2002 to 2010.WV average data value fluctuates more than others', but overall stay the lowest across the states over the year. DC average data value overall stay the highest acrros the states over the year.

##### Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r plot 2006+2010}

plot_2006 = 
  p2_clean %>% 
  filter(location_abbr=="NY",year==2006)|>
  ggplot(aes(x =location_desc , y = data_value,fill=response)) + 
  geom_bar(position="dodge",stat="identity") +  
  labs(title = "Average data value by responses in New York State (2006)") + 
  theme(plot.title = element_text(size = 10), axis.text.x = element_text(angle = 20, hjust = 1),legend.position = "none")

 plot_2010 =
  p2_clean %>%  
  filter(location_abbr == "NY", year == "2010") %>% 
  ggplot(aes(x = location_desc, y = data_value, fill = response)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(title = " Average data value by responses in New York State (2010)") +
  theme(plot.title = element_text(size = 10), axis.text.x = element_text(angle = 20, hjust = 1)) 

options(repr.plot.width = 10, repr.plot.height = 4)
 
 
plot_2006 + plot_2010
```

